MolProbity's Ultimate Rotamer-Library Distributions for Model Validation

### 2016-02-17 response_2_reviwers.txt

"Design libraries and validation libraries focus on two distinct areas of the reference data distributions. While design and prediction are primarily concerned with statistics and cluster shape inside the low-energy wells [Dunbrack 1997], validation is primarily concerned with robustly identifying the outliers beyond the edges of those wells. Such outliers are usually wrong but sometimes valid and interesting, so are always worth examining [Richardson 2013]. “

• Should “chain-level” be expanded in the subheading to the section “Chain-level filters” to “Main-chain” or “backbone conformation”  or “overall quality filters” ?

-->  The subheading has been changed to "Overall Chain-Level Filters", and several small changes have been made in the preceding section to set readers up to realize that "chain" here means a PDB chain, not a polypeptide backbone, and that it is a minor shift from "file-level" filters and contrasts with residue-level filters.

• The following sentence on page 7 made no real sense “This entire set of datasets is therefore nicknamed the ”Top8000” of structural biology’s upper class.“

-->  To make the point clear, we have changed this sentence to read: "We therefore call these datasets the Top8000, as successors to the Top500, now with about 8000 protein chains. This suggests an analogy that the high-resolution, high-quality "upper class" of structural biology has nearly reached the numbers of the "Upper Ten Thousand" or the "Quality" of 19th Century British or American high society [Bisset 1852, Thom 1875]. Quality is less artificial for the Top8000 than for the Upper10000, although they also are not always the ones that accomplish most for the general good."

     We consider this as a small reward and diversion for people who actually read the methods section.

@@

* Section 5.1 should be a single well organized paragraph, not a collection of 5 single sentence paragraphs.  I’m also not completely sure this section (and the designation of ultimate) is completely necessary. It is quite clear you have made a contribution, it is perhaps less clear that we can be absolutely sure we will not want to iterate rotamer library development one or more times after this.

-->  It would be extremely difficult to make this section readable and its logic understandable as a single paragraph.  Instead, we have made it an explicit list of separate arguments and have rewritten it for better clarity and continuity, which is indeed very helpful:

** copy edited text here ***

    The more important issue is the word "ultimate".  Note that even the title says these are MOLPROBITY'S ultimate rotamer distributions (not everyone's).  It is indeed quite likely that someone else will iterate rotamer development -- but we don't intend to, in anything the same form.  We feel the paper makes our case adequately and prominently (in title, abstract, and discussion), and that ultimate can be treated as a meaningful and memorable, but not too dead-serious, designation alongside penultimate.

### 2016-02-17 discussion.tex
Played a little with the wording in 5.1 to make it more reader-frienly

First, we claim it as ultimate for the rotamer aspect of model validation (where outlier status is pre-eminent)  -- such as done by the MolProbity website, in \textit{PHENIX}, or by the wwPDB -- but not necessarily for other purposes (such as protein design) that depend on accurate comparisons within the regions of high probability rather than on accurate delineation of the outer limits of those regions.

Second, changes from the penultimate distributions are fairly minor in quantitative terms;  their importance lies in the new ability to support a 3-level evaluation, to identify very rare rotamers, and to specify rotamer-dependent bond-angle deviations. With a million residues of high-reliability data, the sampling is now adequate for this task. The results are unlikely to change much in the future, except for some special cases such as disulfides or chemically modified amino acids. The multi-dimensional distributions of rotamer probability are freely available on GitHub for use in other software systems.

Third, what we do expect will improve in the future is a change in how sidechain modeling and validation is defined, driven both by expansion of data and compute power and by the needs of structure determination at lower resolution. Rather than separate Ramachandran and rotamer evaluation, we should move toward analyzing all backbone and sidechain torsional dimensions together \cite{JSR_theplot_2013}, including allowance for the influence of secondary structure and local motifs. Thus our "ultimate" claim asserts the position that rotamer evaluation should no longer be updated but should evolve into something else.

>>>>

First, we claim it as ultimate for validation (where outlier status is pre-eminent)  -- such as done by the MolProbity website, in \textit{PHENIX}, or by the wwPDB. We are not advocating for this to be an "ultimate" library for other purposes (such as protein design) that depend on accurate comparisons within the regions of high probability rather than on accurate delineation of the outer limits of those regions.

Second, the quantitative changes from the penultimate distributions are fairly minor. The important new feature is the ability to support a 3-level evaluation to identify very rare rotamers and to specify rotamer-dependent bond-angle deviations. These tasks are made possible due to the large number (a million) of reliably-modeled residues. The results are thus unlikely to change much in the future, except for some special cases not examined here such as disulfides and chemically modified amino acids.

Third, expansion of data and compute power, as well as needs of structure determination at lower resolution, will change how sidechain modeling and validation is defined. Rather than separate Ramachandran and rotamer evaluation, we should move toward analyzing all backbone and sidechain torsional dimensions together \cite{JSR_theplot_2013}, including allowance for the influence of secondary structure and local motifs. Thus our "ultimate" claim asserts the position that rotamer evaluation should no longer be updated but should evolve into something better.
